\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors

### 2022-06-28 Manuscript/Lanuza_et_al_2022.Rmd
edits

Plants have an enormous variation of life history strategies and trait combinations. However, there is strong evidence that there are evolutionary and physiological constraints that limit the number of plant ecological strategies. Despite recent advances in the understanding of the plant trait variation spectrum, reproductive traits are often not considered despite its key role in shaping both plant life-history strategies and plants interactions with floral visitors. Here, we investigate the reproductive spectrum of flowering plants and how this drive the establishment of interactions with their floral visitors using a dataset of 16 reproductive traits for 1,506 plant species. We found that over 50% and 70% of all trait variation was explained by the first two and three reproductive axes, respectively. Specifically, the first axis indicated the presence of a negative correlation between flower number and flower size; the second axis showed a positive correlation between pollinator dependence and floral display; and the third axis indicated a negative correlation between style length and pollinator dependence. These different reproductive axes determined important differences in the interaction level among floral visitor guilds but were insufficient to fully capture plant-pollinator associations. Our study shows the general reproductive trait correlations observed and highlights their relevance to understand plant-pollinator interactions in a global context.

@@

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape their interactions with animal pollinators [@fenster2004; @dellinger2020]. The variety of reproductive strategies followed by different plant species, from specialized reliance on a single pollinator, to full self-pollination, is incredibly rich. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@stearns1989; @agrawal2020]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@salguero2016; @friedman2020; @paterno2020; @roddy2021], their study is often limited to a limited number of reproductive traits that are rarely studied in combination. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other morphological and physiological plant traits such as roots (REF) or leafs (REF) Etc.. [@wright2004; @chave2009; @diaz2016; @onoda2017; @laughlin2021], there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive spectrum can help to further understand the different plant ecological strategies [@agrawal2020] and improve our knowledge on plant-pollinator associations [@roddy2021].

With the recent availability of large trait databases [e.g., TRY @kattge2011;\ COMPADRE @salguero2015], plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, most studies with a multitrait perspective have focused on trait correlations from the leaf [@wright2004], wood [@chave2009], or root [@laughlin2021] related traits with little or no attention given to reproductive traits [@evojtko2020; @roddy2021]. Despite the lack of an holistic view that depict reproductive trait covariation patterns, there are widely recognized reproductive trait associations between pair of traits such as the negative correlation between flower size and flower number [@sargent2007; @kettle2011], the positive association between flower size and outcrossing rate [@goodwillie2010] or the positive association between outcrossing rate and lifespan [@barrett2003; @moeller2017]. Although these different trait correlations (and others) have recently allowed to progress towards a conceptual framework that integrates the different floral trait relationships [@roddy2021], we still lack empirical evidence that investigates jointly how these different reproductive trait associations shape plant life history strategies.

Despite this increasing number of macroecological studies that investigate some plant reproductive traits [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017], we still have poor understanding of how reproductive traits drive interactions with floral visitors at large ecological scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. The use of trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016] has shown to be of great importance when exploring the drivers of plant-pollinator interactions. For example, plant traits can define species' network roles [e.g., specialists vs generalists @tur2013] and plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Indeed, morphological matching between plants and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. However, is still unclear how specific key reproductive traits like mating or compatibility system influence plant-pollinator associations [@tur2013; @devaux2014]. Indeed, the pollination system of a great number of plant species remains unexplored. Because both the species' morphology and  mating system can determine the species' functional role in the pollination network and a combination of traits has shown to increase the predictive [predictive whith respect to what?? what do you want to predict?]() power of the network of interactions [@eklof2013], an interesting novel approach is to investigate how a complete set of reproductive traits in the multidimensional trait space determine species interaction patterns [see @dehling2016 for a morphological approach]. Thus, by exploring the plant reproductive spectrum of trait variation is possible to delimit the different reproductive strategies and explore how these are associated with the different floral visitors.

@@

[Table 1 has a 10^3 which should be 1000]()

@@

T.range <- c("0.01-30m", "0.50-205mm", "0.20-195mm", "0.80-300mm", "0-138mm", "1-25 x $10^3$", "1-18 x $10^4$", "0-160$\\mu$l", "0-6.1mg", "0-100$\\%$", "13-2 x $10^8$", "0-100$\\%$")

@@

First, we investigated how the different groups of floral visitors interacted along the main axes of reproductive trait variation (see below 'visitation patterns' section) with the help of qualitative and quantitative information of plant-pollinator interactions. For this, we used as qualitative information the binary version of the networks (presence-absence of interaction) that assumes equal weight across interactions and as quantitative information the number of visits of floral visitors to individual flowers that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency [@ballantyne2015], the number of visits can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. [Yo aqui meteria los analysis bayesianos. AHora me parece que rompee el flujo]() Second, we investigated how the main axes of trait variation and individual traits influence plant species' functional roles in the pollination network using a set of complementary interaction network metrics: number of visits, normalized degree and specialization (see below 'plant species network roles' section). [Y aqui su analsys]()

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks [or studies?]() but this was accounted for in analyses by considering them in the random effects of the modelling process (see below). All analyses were conducted in R version *4.0.3*. [Y esto es un gran cierre d emetodos]()

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that *Apis mellifera* was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that *A. mellifera* was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of *A. mellifera* on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding *A. mellifera*. We found that *A. mellifera*, was partly driving some of the observed trends on PC1 (figure S5). However, we did not detect major differences on PC2 and PC3.

@@

We investigated whether different quantitative traits determined different plant species' functional roles in the pollination network using Bayesian modelling. For this, we selected simple and complementary species-level network metrics commonly applied in bipartite network studies [@dormann2008] with a straightforward ecological interpretation relevant to our research goals. The different plant species-level metrics were: (i) sum of visits per plant species; (ii) normalized degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) specialization (d') [@bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalization) and 1 (maximum specialization). Normalized degree and specialization were calculated with the *specieslevel* function from the R package *bipartite* [@dormann2008].

We modelled the distinct plant species metrics (sum of visits, normalized degree and plant specialization) as a function of the three main axes of trait variation (plant species level metric \~ PC1 + PC2 + PC3). For each response variable (i.e., each plant species level metric), we used different distribution families (zero inflated negative binomial for the sum of visits, weibull for normalized degree and zero one inflated beta for specialization). Finally, we used the same random factors, model settings and conducted the same posterior predictive checks for each model as detailed above in the 'visitation patterns' section.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, out.width="95%", fig.height=10,fig.align='center',fig.width=10,fig.cap= "\\textbf{Figure 1.} Phylogenetically informed principal component analysis (pPCA) of 1,236 plant species from 28 plant-pollinator network studies. The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) across the two main axes of trait variation. The length of the arrows indicate the weight of the variables on each principal component and the dashed lines show the opposed direction of trait variation. The icons at both ends of arrows and dashed lines illustrate the extreme form of the trait continuum."}

@@

## (c) Visitation patterns 
[asegurate que cuadran los subtitulos de metodos con esto si los modificas]()

@@

The additional model for the most represented families of bees for both presence-absence of interaction (marginal $R^{2}$ = 0.29; conditional $R^{2}$ = 0.19) and number of visits (marginal $R^{2}$ = 0.30; conditional $R^{2}$ = 0.03) showed that the family Apidae was the main driver of the observed patterns. The contrasting differences between presence-absence of interaction and number of visits for bees on PC1 (figure 3a and figure 3d) were driven by the family Andrenidae that had higher number of interacting partners but lower number of visits on plant species with low number of large flowers (figure S9).

@@

The reproductive spectrum of flowering plants show clear positive and negative covariations between reproductive traits that shape the interactions of angiosperms with their floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, the first two and three axes were able to capture over 50% and 70% of reproductive trait variation, respectively. These three axes were: (1) the flower number - flower size, (2) the pollinator dependence - floral display and (3) the style length - pollinator dependence [que opinas de que pol dep salga en dos ejes? No es raro?](). Although the explained trait variation that we found in the first two axes is lower than previous studies that consider only morphological and physiological traits [@diaz2016; @carmona2021], our results are consistent with the largest and most recent study that has characterised plant life strategies including reproductive and plant form and size related traits [@salguero2016]. Interestingly, the different plant reproductive axes were able to partly explain the number of interacting partners of different floral visitor guilds. However, they were insufficient to capture the visitation rates, which highlight the need to account for other local factors, such as local relative abuncances [REF? a malas mi Fun Ecol de 2016](), to better capture the strength of these mutualistic interactions at a macroecological scale.

>>>>

Plants have an enormous variation of life history strategies and trait combinations. However, there is strong evidence that there are evolutionary and physiological constraints that limit the number of plant ecological strategies. Despite recent advances in the understanding of plant trait correlations, reproductive traits are often not considered despite its key role in shaping both plant life-history strategies and plants interactions with floral visitors. Here, we investigate the reproductive spectrum of flowering plants and how this drive the establishment of interactions with their floral visitors using a dataset of 16 reproductive traits for 1,506 plant species. We found that over 50% of all trait variation was explained by the first two reproductive axes. Specifically, the first axis indicated the presence of a negative correlation between flower number and flower size while the second axis showed a positive correlation between pollinator dependence and floral display. These different reproductive axes determined important differences in the interaction level among floral visitor guilds but were insufficient to fully capture plant-pollinator associations. Our study shows the main reproductive trait correlations of flowering plants and highlights their relevance to understand plant-pollinator interactions in a global context.

@@

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape their interactions with animal pollinators [@fenster2004; @dellinger2020]. The variety of reproductive strategies followed by different plant species, from specialized reliance on a single pollinator, to full self-pollination, is incredibly rich. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@stearns1989; @agrawal2020]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@salguero2016; @friedman2020; @paterno2020; @roddy2021], their study is often limited to a few number of reproductive traits that are rarely studied in combination. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other root [@laughlin2021], leaf [@wright2004] or wood [@chave2009] related traits, there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the plant reproductive spectrum can help to further understand the different plant ecological strategies [@agrawal2020] and improve our knowledge on plant-pollinator associations [@roddy2021].

With the recent availability of large trait databases [e.g., TRY @kattge2011;\ COMPADRE @salguero2015], plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, most studies with a multitrait perspective have overlooked plant reproductive traits [@evojtko2020; @roddy2021] and focused on trait correlations from other plant organs. Despite the lack of an holistic view that depict reproductive trait covariation patterns, there are widely recognized reproductive trait associations between pair of traits such as the negative correlation between flower size and flower number [@sargent2007; @kettle2011], the positive association between flower size and outcrossing rate [@goodwillie2010] or the positive association between outcrossing rate and lifespan [@barrett2003; @moeller2017]. Although these different trait correlations (and others) have recently allowed to progress towards a conceptual framework that integrates the different floral trait relationships [@roddy2021], we still lack empirical evidence that investigates jointly how these different reproductive trait associations shape plant life history strategies.

Despite this increasing number of macroecological studies that investigate specific plant reproductive traits [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017], we still have poor understanding of how reproductive traits drive interactions with floral visitors at large ecological scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. The use of trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016] has shown to be of great importance when exploring the drivers of plant-pollinator interactions. For example, plant traits can define species' network roles [e.g., specialists vs generalists @tur2013] and plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Indeed, morphological matching between plants and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. However, is still unclear how specific key reproductive traits like mating or compatibility system influence plant-pollinator associations [@tur2013; @devaux2014]. Indeed, the pollination system of a great number of plant species remains unexplored [@dellinger2020]. Because both the species' morphology and mating system can determine the species' functional role in the pollination network and the combination of several traits has shown to increase the predictive power of the structure of ecological networks [@eklof2013], an interesting novel approach is to investigate how a complete set of reproductive traits in the multidimensional trait space determine species interaction patterns [see @dehling2016 for a morphological approach]. Thus, by exploring the plant reproductive spectrum of trait variation is possible to delimit the different reproductive strategies and explore how these are associated with the different floral visitors.

@@

@@

T.range <- c("0.01-30m", "0.50-205mm", "0.20-195mm", "0.80-300mm", "0-138mm", "1-25000", "1-1.8 x $10^5$", "0-160$\\mu$l", "0-6.1mg", "0-100$\\%$", "13-2 x $10^8$", "0-100$\\%$")

@@

@@

First, we used Bayesian modelling (see below for details) to investigate how the different groups of floral visitors interacted along the main axes of reproductive trait variation with the help of qualitative and quantitative information of plant-pollinator interactions. For this, we used as qualitative information the binary version of the networks (presence-absence of interaction) that assumes equal weight across interactions and as quantitative information the number of visits of floral visitors to individual flowers that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency [@ballantyne2015], the number of visits can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that *Apis mellifera* was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that *A. mellifera* was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of *A. mellifera* on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding *A. mellifera*. We found that *A. mellifera*, was partly driving some of the observed trends on PC1 (figure S5). However, we did not detect major differences on PC2 and PC3.

@@

Second, we investigated with Bayesian modelling how the main axes of trait variation determined different plant species' functional roles in the pollination network. For this, we selected simple and complementary species-level network metrics commonly applied in bipartite network studies [@dormann2008] with a straightforward ecological interpretation relevant to our research goals. The different plant species-level metrics that summarize species' functional roles were: (i) sum of visits per plant species; (ii) normalized degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) specialization (d') [@bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalization) and 1 (maximum specialization). Normalized degree and specialization were calculated with the *specieslevel* function from the R package *bipartite* [@dormann2008]. We modelled the distinct plant species metrics (sum of visits, normalized degree and plant specialization) as a function of the three main axes of trait variation (plant species level metric \~ PC1 + PC2 + PC3). For each response variable (i.e., each plant species level metric), we used different distribution families (zero inflated negative binomial for the sum of visits, weibull for normalized degree and zero one inflated beta for specialization). Finally, we used the same random factors, model settings and conducted the same posterior predictive checks for each model as detailed above in the 'visitation patterns' section.

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across studies but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version *4.0.3*.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, out.width="85%", fig.height=10,fig.align='center',fig.width=10,fig.cap= "\\textbf{Figure 1.} Phylogenetically informed principal component analysis (pPCA) of 1,236 plant species from 28 plant-pollinator network studies. The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) across the two main axes of trait variation. The length of the arrows indicate the weight of the variables on each principal component and the dashed lines show the opposed direction of trait variation. The icons at both ends of arrows and dashed lines illustrate the extreme form of the trait continuum."}

@@

## (c) Visitation patterns

@@

The additional model for the most represented families of bees for both presence-absence of interaction (marginal $R^{2}$ = 0.29; conditional $R^{2}$ = 0.19) and number of visits (marginal $R^{2}$ = 0.30; conditional $R^{2}$ = 0.03) showed that the family Apidae was the main driver of the observed patterns. The contrasting differences between presence-absence of interaction and number of visits for bees on PC1 (figure 3a and figure 3d) were driven by the family Andrenidae that had higher number of interacting partners but lower number of visits on plant species with low number of large flowers (figure S8).

@@

The reproductive spectrum of flowering plants show clear positive and negative covariations between reproductive traits that shape the interactions of Angiosperms with their floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, the first two and three axes were able to capture over 50% and 70% of reproductive trait variation, respectively. These three axes were: (1) the flower number - flower size, (2) the pollinator dependence - floral display and (3) the style length - pollinator dependence [que opinas de que pol dep salga en dos ejes? No es raro?](). Although the explained trait variation that we found in the first two axes is lower than previous studies that consider only morphological and physiological traits [@diaz2016; @carmona2021], our results are consistent with the largest and most recent study that has characterised plant life strategies including reproductive and plant form and size related traits [@salguero2016]. Interestingly, the different plant reproductive axes were able to partly explain the number of interacting partners of different floral visitor guilds. However, they were insufficient to capture the visitation rates, which highlight the need to account for other local factors, such as local relative abuncances [@bartomeus2016], to better capture the strength of these mutualistic interactions at a macroecological scale.
